### 2025-10-03 teaching/math242f25/tut4.tex
typo on page

\proclaim Problem 4. Let $x$ be a real number. Prove that for every $\eps>0$
there exist two rational numbers $q$ and $q'$ such that $q < x < q'$
and $|q-q'|<\eps$.

\proof We know from class that $\QQ$ is dense in $\RR$. We can apply this with
$x-\eps/2$ and $x$ to get $q\in \QQ$ with
$$ x - {\eps \over 2} < q < x.$$
Then we can apply it with $x$ and $x + \eps/2$ to get $q'\in \QQ$ with
$$ x< q' < x + {\eps \over 2}.$$
By construction, we have $q < x < q'$, as desired. It remains to show that
$ |q - q'| < \eps$. Well, we know that $q' < x+\eps/2$ and $q > x-\eps/2$.
Negating the latter, we get $-q < -x + \eps/2$, and adding this to the former,
we obtain
$$|q-q'| = q' - q < {\eps\over 2} + {\eps \over 2} = \eps.$$
So we are done.\slug

>>>>

% \proclaim Problem 4. Let $x$ be a real number. Prove that for every $\eps>0$
% there exist two rational numbers $q$ and $q'$ such that $q < x < q'$
% and $|q-q'|<\eps$.
% 
% \proof We know from class that $\QQ$ is dense in $\RR$. We can apply this with
% $x-\eps/2$ and $x$ to get $q\in \QQ$ with
% $$ x - {\eps \over 2} < q < x.$$
% Then we can apply it with $x$ and $x + \eps/2$ to get $q'\in \QQ$ with
% $$ x< q' < x + {\eps \over 2}.$$
% By construction, we have $q < x < q'$, as desired. It remains to show that
% $ |q - q'| < \eps$. Well, we know that $q' < x+\eps/2$ and $q > x-\eps/2$.
% Negating the latter, we get $-q < -x + \eps/2$, and adding this to the former,
% we obtain
% $$|q-q'| = q' - q < {\eps\over 2} + {\eps \over 2} = \eps.$$
% So we are done.\slug
